---
title: "DATA 607 Final Project: House prices prediction"
author: "Jered Ataky & Zhouzin shi"
date: "2020-11-29"
output: 
  openintro::lab_report: default
  html_document:
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=FALSE}
# Load libraries
library(tidyverse)
library(plyr)
library(kableExtra)
library(plotly)
library(corrplot)
library(PerformanceAnalytics)
library(stats)
library(xgboost)
```

## Problem Statement and background

<style>
div.aquamarine { background-color:#7fffd4; border-radius: 10px; padding: 5px;}
</style>
<div class = "aquamarine">

One of the big decisions to make in life is purchasing a home. There are many factors that need to be taken into consideration while making such a decision. In this project, we tend to retrieve data from different data sources. The data will have metrics such as median income, unemployment rate, public schools, hospitals, hospital ratings, crime rate, …
Our task will be to get data from those multiple data sources using different methods (read csv, web scrapping…) learned throughout the course of this class, to store them (on a database, cloud,…), to clean and transform them, and to analyze and visualize them to get some useful information for houses price.
We will then go further on using different models to predict the house price based on different features that we would find are necessary and weight on houses price.

Research questions : "Which variable is a best predictor of housing prices?";
"What is the relationship between housing prices and each predictor?"

</div> \hfill\break

## Work process 

```{r hide}
Process <- c('Data Collection',
             'Data Transformation (Cleaning & Tidying data)',
             'Data Analysis', 'Visualization',
             'Modeling','Review & Conclusion', 'Presentation')
Team <- c('Jered & Zhouxin', 'Jered & Zhouxin', 'Jered ', 
          'Jered', 'Jered', 'Jered & Zhouxin',
          'Jered & Zhouxin')
df_team <- data.frame(Process, Team)
names(df_team) <- c('Process', 'Team Members')

df_team %>%
  kbl(caption = "Work Process & Responsabilities") %>%
  kable_material(c("striped", "hover")) %>%
  row_spec(0, color = "indigo")
```


## Data

```{r}
# Load the data from Github and GCP storage
hospitals <- read.csv("https://raw.githubusercontent.com/szx868/FinalProject/master/Hospitals.csv")
hospital_ratings <- read.csv("https://raw.githubusercontent.com/szx868/FinalProject/master/Hospital_General_Information.csv")
county_time_series <- read.csv("https://storage.googleapis.com/triplej_project3/County_time_series.csv")
crosswalk = read.csv("https://raw.githubusercontent.com/szx868/FinalProject/master/CountyCrossWalk_Zillow.csv")
unemployment = read.csv("https://raw.githubusercontent.com/szx868/FinalProject/master/Unemployment_Rate_by_County_Percent.csv")
public_schools <- read.csv("https://storage.googleapis.com/triplej_project3/Public_Schools.csv")
```


## Data Transformation

## Tidying the data

```{r}
# Rename columns of unemployment
colnames(unemployment)[4:22] <- seq(2000,2018)
```


```{r}
# Subset hospital_rating data set
hospital_ratings <-hospital_ratings %>% select('Hospital.Name', 'Hospital.overall.rating')
```


```{r}
# Merge hospitals and hospital ratings
hospitals_with_ratings <- merge(hospitals, hospital_ratings, by.x="NAME",by.y = "Hospital.Name")
```


```{r}
# Clean the missing values and subset the hospital data set
hospital <- filter(hospitals_with_ratings, COUNTYFIPS != 'NOT AVAILABLE')
hospital <- filter(hospitals_with_ratings, Hospital.overall.rating != 'Not Available')
```



```{r}
# Rename the columns
hospital <- rename(hospitals_with_ratings,c('FIPS' = 'COUNTYFIPS',  'AverageHospitalRating' = 'Hospital.overall.rating'))
```


```{r}
# Group unemployment per county
unemployment_per_county <- unemployment %>% 
    select('Region.Code', '2018')
unemployment_per_county <- rename(unemployment_per_county, c('FIPS' = 'Region.Code', 'UnemploymentRate' = '2018'))
# Rename unemployment
unemployments <- unemployment_per_county
```


```{r}
# Group public school per county
public_schools_per_county<-public_schools %>%
    group_by(COUNTYFIPS) %>%
    dplyr::summarise(count=n())
```


```{r}
# Subset public schools per county and rename the columns
public_schools_per_county <- public_schools_per_county %>% 
    select('COUNTYFIPS', 'count')
public_schools_per_county <- rename(public_schools_per_county,c('FIPS' = 'COUNTYFIPS', 'NumberOfSchools' = 'count'))
# Rename public_schools_per_county
schools <- public_schools_per_county
```


```{r}
# Subset hospital data set & get hospital average rating
hospital_avg_rating  <- hospital %>% 
    select('FIPS', 'AverageHospitalRating')
```

```{r}
# Convert to numeric
hospital_avg_rating$AverageHospitalRating <- as.numeric(hospital_avg_rating$AverageHospitalRating)
```

```{r}
# Group hospital avg rating by county
hospital_avg <- hospital_avg_rating %>%
    group_by(FIPS) %>%
    dplyr::summarize(AverageHospitalRating = mean(AverageHospitalRating, na.rm=TRUE))
```


```{r}
# Group hospital data set by county
hospitals_per_county <-hospitals %>%
    group_by(COUNTYFIPS) %>%
    dplyr::summarise(count=n())
```


```{r}
# Rename columns of hospital data set
hospitals_per_county <- rename(hospitals_per_county,c('FIPS' = 'COUNTYFIPS', 'NumberOfHospitals' = 'count'))
```


```{r}
# Merge hospitals per county and hospital_avg data set
hospitals_per_county  <- merge(hospitals_per_county,hospital_avg , by.x="FIPS",by.y = "FIPS")
head(hospitals_per_county)
```


```{r}
# Explore housing price from county time series data set
house_prices  <-
county_time_series %>%
    group_by(RegionName) %>%
    dplyr::summarize(ZHVI_AllHomes = mean(ZHVI_AllHomes, na.rm=TRUE))
```


### Put all data set together to form one file

```{r}
# Associate house price with county code
data <- rename(house_prices,c('FIPS' = 'RegionName', 'AverageHousePrice' = 'ZHVI_AllHomes'))
```


```{r}
# Merge crosswalk
data  <- merge(data, crosswalk, by.x="FIPS",by.y = "FIPS")
```


```{r}
# Merge data with unemployment
data  <- merge(data, unemployments, by.x="FIPS",by.y = "FIPS")
```


```{r}
# Merge data with public schools
data  <- merge(data, schools , by.x="FIPS",by.y = "FIPS")
```


```{r}
# Merge data with hospitals_per_county
data  <- merge(data, hospitals_per_county , by.x="FIPS",by.y = "FIPS")
```


```{r}
head(data)
```

```{r}
# Save data as csv
write.csv(data,"data_raw_final.csv", row.names = FALSE)
```

### Cleaning the data

```{r}
# Get raw data from Github
data_raw <- read.csv("https://raw.githubusercontent.com/szx868/FinalProject/master/data_raw_final.csv")
head(data_raw)
```

```{r}
# Drop rows with missing Average house price
data_clean <- data_raw %>%
    drop_na(AverageHousePrice)
# Move the target in the end
data_final <- data_clean %>%
    select(-AverageHousePrice, AverageHousePrice)
head(data_final)
```


## Exploratory Data Analysis

### Which state can you effort living ?

We are going to calculate the average county home price per state.

```{r}
data_1 <- data_final %>%
    group_by(StateName) %>%
    transmute(StateName, avg_house_price = mean(AverageHousePrice))
data_1 <- data_1 %>%
    distinct(StateName, avg_house_price)
data_1
```


**Now we are going to order the state from the least affordable to the most affordable**

```{r}
data_2 <- data_1 %>%
    arrange(desc(avg_house_price))
data_2 %>%
    kbl(caption = "Home price per state") %>%
    kable_material(c("striped", "hover")) %>%
    row_spec(0, color = "indigo")
```

### Visualization

**Visualize the least affordable states based on average county home price per state**

```{r}
# Top 10
top_n(ungroup(data_2), 10) %>%
    ggplot(aes(reorder(StateName, avg_house_price), avg_house_price)) +
    geom_col(aes(fill = avg_house_price)) +
    
    coord_flip() +
    
    labs(title = '10 most expensive state to buy a house', x = "State")
```


**Visualize the most affordable states based on average county home price per state**

```{r}
# Top 10
top_n(ungroup(data_2), -10) %>%
    ggplot(aes(reorder(StateName, avg_house_price), avg_house_price)) +
    geom_bar(stat="identity", color="blue", fill="purple") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=0.5)) +
    
    
    labs(title = '10 least expensive state to buy a house', x = "State")
```

**Visualize the average house price by state**

```{r}
# Add state abbreviation 
data_3 <- data_1 %>%
    mutate(code = state.abb[match(StateName,  state.name)])
# Plot the map
w <- list(color = toRGB("white"), width = 2)
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)
p <- plot_geo(data_3, locationmode = 'USA-states') %>%
  add_trace(
    z = ~avg_house_price, locations = ~code,
    color = ~avg_house_price, colors = 'Purples'
  ) %>%
  colorbar(title = "Avg house price") %>%
  layout(
    title = 'Avg house price by State',
    geo = g
  )
p
```


## Modeling

<style>
div.aquamarine { background-color:#7fffd4; border-radius: 10px; padding: 5px;}
</style>
<div class = "aquamarine">

For this predictive analysis, we are going to use Linear Regression.

Since wee have many explanatory variables, this case will be a multiple linear regression model.

The explanatory variables or predictors are: unemployment rate, number of schools, number of hospitals, average hospital ratings.
Our response variable is the average house price.

Our research question is "which variable is a best predictor of average house price?"

We are going then to test around the relationship between housing prices and each one of the predictors.
The model selection will be based on adjusted R square. Thus, we are going to apply "backward-selection".

The general idea behind backward-selection is to start with the full model and eliminate one variable at a time until the ideal model is reached: Start with the full model, refit all possible models omitting one variable at a time, and choose the model with the highest adjusted R squared, repeat until maximum possible adjusted R squared is reached.

</div> \hfill\break

**Subset the data set with numeric variables to get it ready for modeling**

```{r}
# Subset the data set with only numerical variables
data_4 <- data_final %>%
    group_by(StateName) %>%
    transmute(StateName,
              UnemploymentRate = round(mean(UnemploymentRate), 0),
              NumberOfSchools = sum(NumberOfSchools),
              NumberOfHospitals = sum(NumberOfHospitals),
              AverageHospitalRating = round(mean(AverageHospitalRating), 0),
              avg_house_price = mean(AverageHousePrice))
data_4 <- data_4 %>%
    distinct(StateName, UnemploymentRate, NumberOfSchools, NumberOfHospitals, AverageHospitalRating, avg_house_price)
data_4 <- subset(data_4, select = -c(StateName))
head(data_4)
```


**Correlation matrix**

```{r}
res <- cor(data_4)
round(res, 2)
```


**Performance analytics**


```{r}
data_4 %>%
    chart.Correlation(histogram=TRUE, pch=19)
```

**Summary table of correlation between predictors and house price**

```{r}
Features <- c('NumberOfSchools', 'NumberOfHospitals', 'AverageHospitalRating', 'UnemploymentRate')
Correlation <- c(0.12, -0.02, -0.22, 0.07)
df <- data.frame(Features, Correlation)
df
```


**Multiple linear model:**

We are going to evaluate the avg_house_price with each of the predictors

Since we are using backward-selection, let first start with full model:

```{r}
res_mul <- lm(avg_house_price ~  NumberOfSchools + NumberOfHospitals + AverageHospitalRating + UnemploymentRate, data = data_4)
summary(res_mul)
```


Now, let analyze each individual predictor with house price:

**avg_house_price~UnemploymentRate**

```{r}
res_1 <- lm(avg_house_price ~ UnemploymentRate, data = data_4)
summary(res_1)
```


**avg_house_price~NumberOfSchools**

```{r}
res_2 <- lm(avg_house_price ~ NumberOfSchools, data = data_4)
summary(res_2)
```


**avg_house_price~NumberOfHospitals**

```{r}
res_3 <- lm(avg_house_price ~ NumberOfHospitals, data = data_4)
summary(res_3)
```


**avg_house_price~AverageHospitalRating**

```{r}
res_4 <- lm(avg_house_price ~ AverageHospitalRating, data = data_4)
summary(res_4)
```

**Summary table**

```{r}
Features <- c('NumberOfSchools', 'NumberOfHospitals', 'AverageHospitalRating', 'UnemploymentRate')
Correlation <- c(0.12, -0.02, -0.22, 0.07)
P_values <- c(0.00241, 0.00347, 0.09069,  0.68095)
Adj_r_square <- c(-0.006934, -0.02049,  0.02919, -0.01543)
df_final <- data.frame(Features, Correlation, P_values, Adj_r_square)
df_final
```

## Conclusion

<style>
div.aquamarine { background-color:#7fffd4; border-radius: 10px; padding: 5px;}
</style>
<div class = "aquamarine">

As to answer to our main question to test around relationship between housing prices and each predictors, We realize that the house price has a positive relationship with the number of schools and 
the unemployment rate. Those are also the two big factors (from the features we explored) that contribute the most to the price of home in US (Although the correlation is not strong). The two other factors have a negative relationship with house prices. Though the predictors we used, we should have taken into consideration the crime rate which we believe should be a great predictor of home price. This will be part of further work we will have to do to make this model more efficient. We need also to mention that the project presented some challenges such as we needed to find appropriate data set for various factors, understand different terms such FIPS which we never heard before, merge different data set to make one useful data set for analysis and prediction.

</div> \hfill\break

## References

CUNY DATA606: https://fall2020.data606.net/chapters/chapter9/
